SNP Analysis

### 2019-11-20 snp/SNP_Prediction_Analysis.Rmd
bug fix, forgot to output sample names in the tiled files

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(gridExtra)
library(glmnet)
library(pROC)
#library(ggfortify)
library(BarrettsProgressionRisk)

@@

@@

qcdata = qcdata %>% as_tibble %>% 
  separate(Samplename, c('PatientID','SampleID','EndoID','Level'), sep='_', remove=F)

@@

qcdata[1,] -> s

qcdata$ASCAT.SCA.ratio = apply(qcdata,1,function(s) {
  smp = filter( segments.list[[ s$PatientID ]], sample == s$Samplename & chr %in% c(1:22) )
  smp = smp %>% #rowwise %>% 
    dplyr::mutate(
      #'Total' = nAraw + nBraw,
      'Total' = nMajor + nMinor,
      'CNV' = round(Total) - round(as.numeric(s$Ploidy)) )

  x = filter(smp, CNV != 0 & chr %in% c(1:22))
  sum(x$endpos - x$startpos)/as.numeric(chr.info[22,'genome.length'])
})

qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',qcdata$Level, ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',qcdata$Level, ignore.case=T)] = 'Gastric Normal'

patient.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status) %>% 
  mutate(Path.Status = as.character(ifelse(Pathology %in% c('IMC','HGD'), 'P', Path.Status)))

sample.list = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
colnames(sample.list)[3] = 'Total.SCA'

sample.list = sample.list %>% mutate(SCA.Ratio = Total.SCA/max(chr.info$genome.length))

sample.info = qcdata[,c('PatientID','SampleID','EndoID','Level','Samplename','ASCAT.SCA.ratio')]

@@

sample.info = base::merge(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by.x=c('PatientID','EndoID'), by.y=c('PatientID','Timepoint Code'))
head(sample.info)

@@

# ASCAT and Xiahong's estimates correlate in 86% of samples, that suggests that the 14% missing is due to the LOH I'm not trying to call
cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(gridExtra)
library(glmnet)
library(pROC)
#library(ggfortify)
library(BarrettsProgressionRisk)
library(kableExtra)

@@

patient.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status, PatientID = as.character(PatientID)) %>% 
  mutate(Path.Status = as.character(ifelse(Pathology %in% c('IMC','HGD'), 'P', Path.Status)))

patient.info %>% filter(PatientID == 524)

@@

qcdata = qcdata %>% as_tibble %>%  mutate(Samplename = sub('_$','',Samplename)) %>%
  separate(Samplename, c('PatientID','SampleID','EndoID','Level'), sep='_', remove=F)

qcdata[749, 'Level'] = 'BLD' # Making an assumption here but this is what it looks like

@@

calc.ratio<-function(PatientID, Samplename,Ploidy) {
  smp = filter( segments.list[[ PatientID ]], sample == Samplename & chr %in% c(1:22) )
  smp = smp %>% dplyr::mutate(
      #'Total' = nAraw + nBraw,
      'Total' = nMajor + nMinor,
      'CNV' = round(Total) - round(as.numeric(Ploidy)) )
  x = filter(smp, CNV != 0 & chr %in% c(1:22))
  return(sum(x$endpos - x$startpos)/as.numeric(chr.info[22,'genome.length']))
}

qcdata = qcdata %>% mutate(
  ASCAT.SCA.ratio = purrr::pmap_dbl(list(PatientID, Samplename, Ploidy), .f=calc.ratio),
  Level = toupper(Level),
  SampleType = case_when(
    grepl('BLD', Level, ignore.case = T) ~ 'Blood Normal',
    grepl('gastric', Level, ignore.case = T) ~ 'Gastric Normal',
    TRUE ~ 'BE'
))

sample.list = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2) %>%
  dplyr::rename('Total.SCA' = 3) %>% mutate(SCA.Ratio = Total.SCA/max(chr.info$genome.length))

sample.info = dplyr::select(qcdata, c('PatientID','SampleID','EndoID','Level','Samplename','ASCAT.SCA.ratio'))

@@

sample.info = left_join(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by=c('PatientID','EndoID' = 'Timepoint Code')) %>%
  mutate(Status = factor(as.character( ifelse(is.na(Status) & Level %in% c('BLD','GASTRIC'), 'Normal', Status))))
head(sample.info)

@@

ct = cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)
ct$estimate
```

ASCAT and Xiahong's estimates show a `r round(ct$estimate,2)` correlation. That suggests that the remainder is due to the LOH I'm not trying to call.
